---
title: "hw2_xl2851"
author: "xinran"
date: "3/21/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(caret)
library(patchwork)
library(splines)
library(gam)
library(mgcv)
library(boot)
library(ggplot2)
library(pdp)
library(earth)
```

```{r}
#import data
college = read_csv("./data/College.csv")%>%
  janitor::clean_names()
#exclude Columbia University and the "college" column
college_tr = college %>%  
  filter(college != "Columbia University") %>% 
  select(-college)
```

## Set random seed

```{r}
set.seed(1)
```


## (a) Create scatter plots of response vs. predictors.

response vs predictors

```{r,fig.height=12, fig.width=12}
# matrix of predictors
x = model.matrix(outstate ~ .,college_tr)[,-1] 
# vector of response
y = college_tr$outstate

theme1 <- trellis.par.get()
theme1$plot.symbol$col <- rgb(.2, .4, .2, .5) 
theme1$plot.symbol$pch <- 16
theme1$plot.line$col <- rgb(.8, .1, .1, 1) 
theme1$plot.line$lwd <- 2
theme1$strip.background$col <- rgb(.0, .2, .6, .2)
trellis.par.set(theme1)
featurePlot(x, y, plot = "scatter", labels = c("","Y"), type = c("p"), layout = c(4, 4))
```

## (b) Fit a smoothing spline model using Terminal as the only predictor of Outstate for a range of degrees of freedom, as well as the degree of freedom obtained by generalized crossvalidation, and plot the resulting fits. Describe the results obtained.

```{r}
fit.ss=smooth.spline(college_tr$terminal)

fit.ss$df

terminallims <- range(college_tr$terminal)
terminal.grid <- seq(from = terminallims[1],to = terminallims[2])

pred.ss <- predict(fit.ss,
                   x = terminal.grid)

pred.ss.df <- data.frame(pred = pred.ss$y,
                         terminal = terminal.grid)

p <- ggplot(data = college_tr, aes(x = terminal, y = outstate)) +
     geom_point(color = rgb(.2, .4, .2, .5))
p +
geom_line(aes(x = terminal, y = pred), data = pred.ss.df,
          color = rgb(.8, .1, .1, 1)) + theme_bw()
```

description: The degree of freedom generalized by cross validation is `r fit.ss$df` There is an increasing, monotonic but non-linear relationship between terminal and outstate. There seems to be a smooth model that fits the data.

## (c) Fit a generalized additive model (GAM) using all the predictors. Plot the results and explain your findings.
```{r}
ctrl1 <-trainControl(method = "cv", number = 10)
set.seed(1)
gam.fit <-train(x, y,
                method = "gam",
                tuneGrid =data.frame(method = "GCV.Cp", select =c(TRUE,FALSE)),
                trControl = ctrl1)

gam.fit$bestTune
```

```{r}
gam.fit$finalModel
```

```{r}
plot(gam.fit$finalModel)
```

## (d)(1) Fit a multivariate adaptive regression spline (MARS) model using all the predictors. Report the final model. 

```{r}
mars_grid <- expand.grid(degree = 1:2, 
                         nprune = 2:10)

set.seed(1)
mars.fit <- train(x, y,
                 method = "earth",
                 tuneGrid = mars_grid,
                 trControl = ctrl1)

ggplot(mars.fit)

mars.fit$bestTune

```

The final model is:
```{r}
coef(mars.fit$finalModel)
```

## (d)(2)Present the partial dependence plot of an arbitrary predictor in your final model.
```{r}
p1 <- partial(mars.fit, pred.var = c("perc_alumni"), grid.resolution = 10) %>% autoplot()

grid.arrange(p1)
```

## (e) Based on the above GAM and MARS models, predict the out-of-state tuition of Columbia University.
### Predict basing on GAM

```{r}
columbia_u = college %>% 
  filter(college == "Columbia University")
pred.gam = predict(gam.fit,
                   newdata = columbia_u )
pred.gam
```

The out-of-state tuition of Columbia University is predicted to be 17728.51$ by the above GAM model.

### Predict basing on MARS

```{r}
pred.mars = predict(mars.fit,
                    newdata = columbia_u)
pred.mars
```

The predicted out of state tuition for Columbia University using the MARS model is 17469.90$.
